\documentclass[11pt,a4paper]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,booktabs,hyperref,float}

\title{Iteration 018: Euclidean Alignment}
\author{Autoresearch Loop}
\date{April 8, 2026}

\begin{document}
\maketitle

\begin{abstract}
Per-batch whitening via Euclidean alignment achieves $r=0.369$, worse than
the baseline ($r=0.378$). Noisy batch-level covariance estimates at test time
degrade rather than improve predictions.
\end{abstract}

\section{Hypothesis}
Euclidean alignment (He \& Wu, 2019) whitens each subject's data to a common
reference, reducing cross-subject covariance differences. Applied per-batch
at test time, it could adapt the model to each test subject's spatial covariance.

\section{Method}
For each batch, compute the sample covariance matrix $\hat{\Sigma}$, then whiten:
$\tilde{x} = \hat{\Sigma}^{-1/2} x$. Applied to both training and test batches.
Combined with the iter017 training recipe.

\section{Results}
\begin{table}[H]
\centering
\begin{tabular}{lrrr}
\toprule
Model & Mean $r$ & Std $r$ & SNR (dB) \\
\midrule
Combined loss (iter017) & 0.378 & 0.078 & 0.61 \\
Euclidean alignment & 0.369 & 0.072 & 0.60 \\
\bottomrule
\end{tabular}
\end{table}

\section{Analysis}
The $-$0.009 degradation reveals that batch-level covariance estimation is too
noisy with 2s windows at 20\,Hz (only 40 timepoints per window). The covariance
matrix is $27 \times 27$ but estimated from very few samples, leading to unstable
whitening. Subject-level alignment (requiring subject labels) might work, but
per-batch alignment is unreliable at this scale.

\section{Next}
Iteration 019: stochastic weight averaging for flatter minima.

\end{document}
