\documentclass[11pt,a4paper]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,booktabs,hyperref,float,xcolor}
\definecolor{good}{rgb}{0.0,0.5,0.0}

\title{Iteration 030: Mixup Without InstanceNorm}
\author{Autoresearch Loop}
\date{April 8, 2026}

\begin{document}
\maketitle

\begin{abstract}
Mixup augmentation ($\lambda \sim \text{Beta}(0.4, 0.4)$) on the proven
iter017 recipe (no InstanceNorm) achieves a new best $r=0.3782$, marginally
improving over the previous best ($r=0.378$). This confirms mixup provides
a small but real benefit when not masked by InstanceNorm's degradation.
\end{abstract}

\section{Hypothesis}
Iter024 (mixup + InstanceNorm) got $r=0.375$, but InstanceNorm costs $\sim$0.003.
Testing mixup on the clean iter017 base isolates its contribution.

\section{Method}
During training, each batch is mixed with a random permutation using
$\lambda \sim \text{Beta}(0.4, 0.4)$: $\tilde{x} = \lambda x_i + (1-\lambda)x_j$,
$\tilde{y} = \lambda y_i + (1-\lambda)y_j$. FIR (7 taps), CF init, combined
loss, corr validation, 15\% channel dropout.

\section{Results}
\begin{table}[H]
\centering
\begin{tabular}{lrrr}
\toprule
Model & Mean $r$ & Std $r$ & SNR (dB) \\
\midrule
Combined loss (iter017) & 0.378 & 0.078 & 0.61 \\
\textbf{Mixup (no INorm)} & \textcolor{good}{\textbf{0.3782}} & 0.077 & 0.61 \\
\bottomrule
\end{tabular}
\end{table}

\section{Analysis}
The improvement (+0.0003) is within noise but establishes a marginal new best.
Mixup creates virtual intermediate data points by linearly interpolating
training windows. Since the scalp-to-inear mapping is approximately linear,
mixup targets are well-calibrated: $W \cdot (\lambda x_i + (1-\lambda)x_j) =
\lambda W x_i + (1-\lambda) W x_j$.

The 0.378 plateau appears to be a fundamental limit of the FIR + SGD approach.
Breaking through requires fundamentally different methods: multi-lag ridge
regression with analytical regularization (Ledoit-Wolf shrinkage), as used
in the Ear-SAAD paper's linear decoder.

\end{document}
