\documentclass[11pt,a4paper]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,booktabs,hyperref,float}

\title{Iteration 020: InstanceNorm + FIR}
\author{Autoresearch Loop}
\date{April 8, 2026}

\begin{document}
\maketitle

\begin{abstract}
Per-window InstanceNorm before the FIR filter achieves $r=0.375$, slightly
below the best ($r=0.378$). Normalizing each 2s window independently removes
useful amplitude information that the FIR filter exploits for prediction.
\end{abstract}

\section{Hypothesis}
Cross-subject covariance differences are the bottleneck. InstanceNorm normalizes
each window independently per channel, making the model invariant to local
amplitude fluctuations. Unlike trial-level z-scoring in preprocessing, this
normalizes \emph{each window}, removing within-trial amplitude dynamics.

\section{Method}
Architecture: InstanceNorm1d (affine=True) $\to$ FIR (7 taps, CF init).
Training: combined MSE+corr loss, corr validation, 15\% channel dropout,
150 epochs. Same hyperparameters as iter017.

\section{Results}
\begin{table}[H]
\centering
\begin{tabular}{lrrr}
\toprule
Model & Mean $r$ & Std $r$ & SNR (dB) \\
\midrule
Combined loss (iter017) & 0.378 & 0.078 & 0.61 \\
InstanceNorm + FIR & 0.375 & 0.077 & 0.63 \\
\bottomrule
\end{tabular}
\end{table}

\section{Analysis}
InstanceNorm slightly hurts correlation despite improving SNR (+0.02\,dB). The
per-window normalization removes slow amplitude modulations (e.g., alpha power
fluctuations within a trial) that correlate between scalp and in-ear channels.
These correlated amplitude dynamics carry useful information for prediction.

Notably, the higher SNR indicates that InstanceNorm helps with absolute scale
matching, but the lower Pearson $r$ shows it removes temporal structure that
matters for waveform shape prediction.

This result suggests that the preprocessing z-scoring (per trial) is already
near-optimal: it removes cross-subject scale differences while preserving
within-trial dynamics.

\section{Next}
Iteration 021: Mixture of Experts to allow input-adaptive spatial filtering.

\end{document}
